---
title: "Hayman Fire Recovery"
author: "Matthew Ross and Orion Cardenas"
date: "2/2/2022"
output: html_document
---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='..')
```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('data',full.names=T)


#Read in individual data files
ndmi <- read_csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read_csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read_csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```




## Question 1) 

What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation s a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months

There is a positive correlation between NDMI and NDVI. I observe higher moisture and NDVI values in the unburned site, and lower moisture and NDVI values in the burned site. This makes sense provided that adequate moisture is necessary for vegetation growth.

```{r}
full_wide <- spread(data=full_long,key='data',value='value') %>%
  filter_if(is.numeric,all_vars(!is.na(.))) %>%
  mutate(month = month(DateTime),
         year = year(DateTime))

summer_only <- filter(full_wide,month %in% c(6,7,8,9))

ggplot(summer_only,aes(x=ndmi,y=ndvi,color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))
```


## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?
 
There does not appear to be any correlation between the previous year's snow cover and the vegetation growth in the following year. The values appear to be uncorrelated although we still observe higher NDVI values in the unburned site as expected.
``` {r}
avg_ndsi_snow_mos <- full_wide %>%
  mutate(year=year(DateTime)) %>%
  mutate(month=month(DateTime)) %>%
  filter(month %in% c(1,2,3,4)) %>%
  group_by(site,year) %>%
  summarize(mean_ndsi=mean(ndsi))

avg_ndvi_summer <- full_wide %>%
  mutate(year=year(DateTime)) %>%
  mutate(month=month(DateTime)) %>%
  filter(month %in% c(6,7,8)) %>%
  group_by(site,year) %>%
  summarize(mean_ndvi=mean(ndvi))

q2_ndvi_ndsi <- left_join(avg_ndsi_snow_mos, avg_ndvi_summer)

ggplot(q2_ndvi_ndsi,aes(x=mean_ndsi,y=mean_ndvi,color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))
```

## Q3

How is the snow effect from question 2 different between pre- and post-burn
and burned and unburned? 

Pre burn there is little difference in the NDVI and NDSI values across the
burned and unburned sites. Post burn there is an obvious difference between the
sites in which the unburned site has high NDVI and NDSI values and the burned site
has low NDVI and NDSI. 

```{r}

wide_burn_unburned <-spread(data=q2_ndvi_ndsi,key = "site", value = "site")

pre_burn <- q2_ndvi_ndsi %>%
  filter(year < 2002)

post_burn <- q2_ndvi_ndsi %>%
  filter(year > 2002)

ggplot(pre_burn,aes(x=mean_ndsi,y=mean_ndvi, color = site)) + 
  ggtitle("Pre Burn") +
  xlab("NDSI") +
  ylab("NDVI") +
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))

ggplot(post_burn,aes(x=mean_ndsi,y=mean_ndvi, color = site)) + 
  ggtitle("Post Burn") +
  xlab("NDSI") +
  ylab("NDVI") +
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))

```

## Question 4

What month is the greenest month on average? 

August has the highest NDVI on average indicating it is the "greenest".

``` {r}
avg_ndvi_month <- full_wide %>%
  mutate(year=year(DateTime)) %>%
  mutate(month=month(DateTime)) %>%
  group_by(month) %>%
  summarize(mean_ndvi=mean(ndvi))

ggplot(avg_ndvi_month,aes(x=month,y=mean_ndvi)) + 
  ggtitle("Average NDVI By Month") +
  xlab("Month") +
  ylab("NDVI") +
  scale_x_continuous(breaks = seq(0, 12, by = 1)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))
```

## Question 5) 

What month is the snowiest on average?

January has the highest NDSI on average indicating it is the "snowiest".
``` {r}
avg_ndsi_month <- full_wide %>%
  mutate(year=year(DateTime)) %>%
  mutate(month=month(DateTime)) %>%
  group_by(month) %>%
  summarize(mean_ndsi=mean(ndsi))

ggplot(avg_ndsi_month,aes(x=month,y=mean_ndsi)) + 
  ggtitle("Average NDSI By Month") +
  xlab("Month") +
  ylab("NDSI") +
  scale_x_continuous(breaks = seq(1, 12, by = 1)) +
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.2))
```

